---
title: "Tutorial Benthic Community Comparison"
author: "Vianney DENIS"
date: "2026-05-08"
output: html_document
---
# Hello world !

Script + data analysis adpated from Ribas Deulofeu et al. 2016

# **Packages and functions**

```{r eval = TRUE, echo=T, message=F, warning=F}
library(vegan)
library(gclus)
source('Function/coldiss.R')
```

# **Generate R Script**

```{r eval = TRUE,  message=F, warning=F, purl=F, results="hide"}
knitr::purl("Notebook.Rmd", documentation = F)
```


# **Dataset**

Import dataset and remove all unstable substrate (transect level).

```{r eval = TRUE, echo=T, message=F, warning=F}
data<- read.csv('Data/TWdataset.csv',h=TRUE,row.names='OTUs') # import dataset
str(data)
data1<-as.data.frame(data) # conversion as a dataframe
data1[,'sand']<- NULL    #  delete unstable substrate 'sand'
data1[,'gravel_small_rubble']<- NULL #  delete unstable substrate 'gravel_small_rubble'
data1[,'rubble']<- NULL #  delete unstable substrate 'gravel_small_rubble'
data1<-as.matrix(data1) # conversion as a matrix
data2<-prop.table(data1,1) # conversion in percentage
```

Import a **factor** variable

```{r eval = TRUE, echo=T, message=F, warning=F}
fact1<-read.csv('Data/TWfactors.csv', h=TRUE,row.names=1)
head(fact1)
```

Aggregate dataset to site level

```{r eval = TRUE, echo=T, message=F, warning=F}
SiteData<- aggregate(data2, by=list(Category=fact1$site), FUN=mean) # mean OTUs percentages per sites
rownames(SiteData)<-SiteData$Category # apply Category (Site) as rownames
SiteData[,'Category']<- NULL    #  delete 'Category' column as variable 
```

Create a new **factor** variable: region (site level)

```{r eval = TRUE, echo=T, message=F, warning=F}
SiteName<- c('Bitou','Chaikou','Chinwan_Inner_Bay', 'Cimei', 'Dabaisha','Dingbaisha',  'Gongguan', 'Gupoyu','Hongchai', 'Houbihu', 'Jialeshuei', 'Keelung_Island','Leidashih', 'Longdon','Longkeng', 'Outlet', 'Pon_Pon_Tan','Sangjiaowan', 'Shihland', 'Siyuping','Tiaoshih',  'Tanzihwan','Wa_En_Tung', 'Wanlitung','Yeliu')
RegionName<- c('North_Taiwan','Green_Island','Penghu','Penghu','Green_Island','Kenting','Green_Island','Penghu','Kenting','Kenting','Kenting','North_Taiwan','Kenting','North_Taiwan','Kenting','Kenting','Penghu','Kenting','Green_Island','Penghu','Kenting','Kenting','Penghu','Kenting','North_Taiwan')
fact1<-data.frame(SiteName,RegionName) # regional factor at site level
```

# **Data analysis**

Distance matrix 

```{r eval = TRUE, echo=T, message=F, warning=F}
## distance & cluster##
dis1<- vegdist(SiteData, method='bray') # bray-curtis similarity (site level)
coldiss(dis1,byrank=F,diag=T) # for the bc dissimilarity on raw data 
```

Cluster

```{r eval = TRUE, echo=T, message=F, warning=F}
clus1<-hclust(dis1, method='average') # cluster average linkage
plot(clus1)
```

## nonmetric Multidimensional Scaling ##

Create an empty plot and add elements one by one

```{r eval = TRUE, echo=T, message=F, warning=F}
mds1site <- metaMDS(SiteData, dist='bray', try=99,type='n') # nonmetric Multidimensional Scaling analysis 
plot(mds1site, display='sites', type='none') # plot Nonmetric Multidimensional Scaling
text(mds1site, 'species',col = rgb(0,0,0,alpha=0.3) , cex=0.5) # visualize species names
points(mds1site, 'sites', pch=19,cex=2, col='yellow',select=fact1$RegionName=='Penghu') # add transects from Penghu
points(mds1site, 'sites', pch=19, cex=2, col='green',select=fact1$RegionName=='Green_Island')# add transects from Green Island
points(mds1site, 'sites', pch=19, cex=2,col='cyan',select=fact1$RegionName=='North_Taiwan') # add transects from North Taiwan
points(mds1site, 'sites', pch=19, cex=2,col='red',select=fact1$RegionName=='Kenting') # add transects from Kenting
ordispider (mds1site, RegionName, col='red') # position centroids
```

Simper test

```{r eval = TRUE, echo=T, message=F, warning=F}
simper1site<-simper(SiteData, RegionName) # similarity percentages analysis
simper1site # discriminating species between regions
```

Dispersion test

```{r eval = TRUE, echo=T, message=F, warning=F}
mod<-betadisper (dis1,RegionName) # multivariate homogeneity of groups dispersions on untransformed data (site Level)  
permutest (mod, pairwise=T) # ANOVA like permutation test (see ?permutest)
TukeyHSD(mod) # pairwise comparison
boxplot(mod) # visualization of the dispersion
```

Test of regional difference using permanova

```{r eval = TRUE, echo=T, message=F, warning=F}
permanovaRegion<-adonis2(formula=SiteData~RegionName, data=fact1, permutations=9999, method='bray') # Permutational Multivariate Analysis of Variance using bray-curtis  
permanovaRegion # permanova results
```

Test of regional difference using anosim

```{r eval = TRUE, echo=T, message=F, warning=F}
anosim1<-anosim(dis1,RegionName, permutations=9999) # analysis of similarities
anosim1 # results anosim
```


